---
title: 'Data Challenge #2'
authors: "Brandon Wolff, Zachary Heinemann, and Stephanie Langeland"
due date: "3/22/2017"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a team assignment. Please create a file on your team GitHub repo where 
you answer the challenge, including links to your code, and graphs.

Group notes (to be deleted):
  * should we all do one model on our own each, then we all do the 4th model collectively to evenly spread out the work?
  
  * did he add more data/variables?
  
  *use the chucks is needed, delete the chuck under each question if its not needed.
  
#__Model #1 for question #1:__

###1.	Ask two (2) questions that might help you understand better the dynamics of 
###violence contained on our datas et. Apply one algorithm per question and share 
###your insights from each analysis. [50 pts] Remember: a non-finding is also a 
###finding! It tells you whether a question is worth pursuing further or not.
  *
   
```{r}

```

###1a	perform the necessary transformations in your data - if any are needed, 
###and explain why you did that
  * 
  
```{r}

```

###1b	show the output from your analysis in a consumable form
  * 

```{r}

```

###1c	be explicit about the limitations of your anaylisis, due to estimation 
###or to the data itself
  * 
  
```{r}

```

###1d	did you find something interesting? what is that? does your finding 
###suggest this question is worth pursuing further? why or why not?
    * 
  
```{r}

```

###1e	if you did not find something interesting, explain why, and whether 
###there is some additional information that would help in answering your question
  * 
  
```{r}

```

###1f	provide your code, and a single visualization per question that summarizes 
###your finding
  * 
  
```{r}

```

###1g	phrase your finding for each question in two ways:
###1g-1	one sentence that summarizes your insight
  * 

###1g-2	one paragraph that reflects all nuance in your insight
  * 
        
###1h	make sure to also include your code
  * The code is included under each quetsion above.
  
#__Model #2 for question #1:__

###1.	Ask two (2) questions that might help you understand better the dynamics of 
###violence contained on our datas et. Apply one algorithm per question and share 
###your insights from each analysis. [50 pts] Remember: a non-finding is also a 
###finding! It tells you whether a question is worth pursuing further or not.
  *
   
```{r}

```

###1a	perform the necessary transformations in your data - if any are needed, 
###and explain why you did that
  * 
  
```{r}

```

###1b	show the output from your analysis in a consumable form
  * 

```{r}

```

###1c	be explicit about the limitations of your anaylisis, due to estimation 
###or to the data itself
  * 
  
```{r}

```

###1d	did you find something interesting? what is that? does your finding 
###suggest this question is worth pursuing further? why or why not?
    * 
  
```{r}

```

###1e	if you did not find something interesting, explain why, and whether 
###there is some additional information that would help in answering your question
  * 
  
```{r}

```

###1f	provide your code, and a single visualization per question that summarizes 
###your finding
  * 
  
```{r}

```

###1g	phrase your finding for each question in two ways:
###1g-1	one sentence that summarizes your insight
  * 

###1g-2	one paragraph that reflects all nuance in your insight
  * 
        
###1h	make sure to also include your code
  * The code is included under each quetsion above.
  
#__Model #1 for question #2:__
  
###2.	Formulate two (2) conditional hypotheses that you seek to investigate with 
###the data. One of your hypotheses should condition on two variables (as the 
###example on the slides), and the other should condition on three variables. [50 pts]
    * __conditional on 2 variables___:  Stephanie: here people = civilians. given that the army 
    was involved, how many cartridges were seized, how many people were wounded, and how many 
    people died? X1 = civilians wounded, X2 = civilians dead, and Y = cartidges seized.  We know 
    the army was there how?  did he give us this info?  When the army is involved, the more 
    cartriges to be sezied are wounded and dead increase?  its easier to take cartidges from 
    dead and wounded people than those still actively fighting. Also, we are assuming that the 
    army is bringing more serious weapons so more cartridges may have been seized and they are 
    more effective than the other forces so fewer people will be wounded. Use a regression model
    and look at the traditional stats like the ones on the slides from 03/01/17

```{r}

```

  
###2a	formulate each one of your hypotheses explicitly in substantive terms (as 
###opposed to statistical terms) using 2-3 lines at most
  * 
  
###2b	show exactly how each one of your hypotheses translates into the marginal 
###effect that you will seek to estimate from the data
  * 
  
###2c	show the output from your analysis in a consumable form
  * 

```{r}

```

###2d	show all your computations to estimate the corresponding marginal effect 
###and its standard error
  * 

```{r}

```

###2e	be explicit in your assumptions
  * 
  
###2f	be explicit in the limitations of your inferences
  * 
  
###2g	phrase your finding for each question in two ways:
###2g-1	one sentence that summarizes your insight
  * 
  
###2g-2	one paragraph that reflects all nuance in your insight
  * 
  
###2h	make sure to also include your code
  * The code is included under each quetsion above.

#__Model #1 for question #2:__

###2.	Formulate two (2) conditional hypotheses that you seek to investigate with 
###the data. One of your hypotheses should condition on two variables (as the 
###example on the slides), and the other should condition on three variables. [50 pts]
    * __conditional on 3 variables___:  

```{r}

```
  
###2a	formulate each one of your hypotheses explicitly in substantive terms (as 
###opposed to statistical terms) using 2-3 lines at most
  * 
  
###2b	show exactly how each one of your hypotheses translates into the marginal 
###effect that you will seek to estimate from the data
  * 
  
###2c	show the output from your analysis in a consumable form
  * 

```{r}

```

###2d	show all your computations to estimate the corresponding marginal effect 
###and its standard error
  * 

```{r}

```

###2e	be explicit in your assumptions
  * 
  
###2f	be explicit in the limitations of your inferences
  * 
  
###2g	phrase your finding for each question in two ways:
###2g-1	one sentence that summarizes your insight
  * 
  
###2g-2	one paragraph that reflects all nuance in your insight
  * 
  
###2h	make sure to also include your code
  * The code is included under each quetsion above.