# #######################################################################
# File-Name:  demographics and crime stats_v1
# Version:  1
# Date: 03/04/17 
# Author: Stephanie Langeland 
# Purpose: Find demographic and crime stats data/clean the data
# Input Files: 2015_stopandfrisk_CLEAN_w_counties.csv, crime data_combined_raw2.csv
# Output Files: cleaned_2015_crime_data.csv
# Data Output: None
# Dependencies: None
# Required by: Final Prject
# Status: In progress 
# Machine: Stephanie's 2011 MacBook Pro
# R version: 3.3.1
# #######################################################################

# Stop and Frisk Data: General:
```{r}
rm(list = ls(all = TRUE))   # cleans everything in the workspace

sfd <- read.csv("/Users/StephanieLangeland/Desktop/Columbia/Applied Data Science/Git/QMSS_G5069_Applied_D_S/Data:Code Book/Cleaned/2015_stopandfrisk_CLEAN_w_counties.csv") # cleaned stop and frisk data 

# SPACE FOR SOMEONE ELSE TO COMMENT OUT MY PATH ABOVE AND PUT THEIR PATH IN HERE
```

# Stop and Frisk Data: Time period of the data:
```{r}
# convert to date format:
library(lubridate)

summary(sfd$datestop)
summary(is.na(sfd$datestop))
typeof(sfd$datestop)
sfd$datestop <- mdy(sfd[ , 5]) # column 5 is the DATE OF STOP (M-D-YYYY) <- the code book is wrong, it says that the format is (MM-DD-YYY)
summary(is.na(sfd$datestop))

min(sfd$datestop) #confirmed that the data are complete for 2015
max(sfd$datestop)

summary((sfd$dob)) # SUSPECT'S DATE OF BIRTH (CCYY-MM-DD) is all missing -> use `age` variable instead
```

# Demographic data for the Bronx: General 

data source: https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF 
```{r}
#the raw dataset was such a mess that I had to clean it in excel - the follwing is the cleaned version:


# this dataest is in a really strange format 
pop <- read.csv("/Users/StephanieLangeland/Desktop/Columbia/Applied Data Science/Git/QMSS_G5069_Applied_D_S/Data:Code Book/Cleaned/CLEAN_ACS_15_5YR_DP05.csv")

#pop <- pop[-22, ] # delete duplicate rows
#pop <- pop[-21, ] # delete duplicate rows

# Using the "tp_Race alone or in combination with one or more other races" category:
library(tibble)

demographic <- data.frame(subset(pop[53:58, 1:2]))

typeof(demographic$Estimate)
str(demographic)
demographic$Estimate <- as.numeric(gsub(",", "", as.character(demographic$Estimate)))
demographic

demographic[ , 1]
demographic$Subject <- c("White",
              "Black or African American",
              "American Indian and Alaska Native",
              "Asian",
              "Native Hawaiian and Other Pacific Islander",
              "Some other race")

library(ggplot2)
ggplot(demographic, aes(x = Subject, y = Estimate)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Compare stop and frisk events to demographic data:

```{r}
ggplot(demographic, aes(x = Subject, y = Estimate)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Race") +
  ylab("Population Estimate") +
  ggtitle("2015 Bronx Population by Race") +
  geom_text(aes(label = Estimate), vjust = 0, colour = "black")

stops_by_race <- sfd[ , "race"]

stops_by_race2 <- summary(stops_by_race)
stops_by_race2 <- as.data.frame(stops_by_race2)
stops_by_race2
stops_by_race2$race <- rownames(stops_by_race2)
stops_by_race2$count <- stops_by_race2$stops_by_race2

ggplot(stops_by_race2, aes(x = race, y = count)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Number of Stop & Frisk Events") +
  xlab("Race") +
  ggtitle("Stop & Frisk Events by Race") +
  geom_text(aes(label = count), vjust = 0, colour = "black")
```

#FINAL GRAPHS for 03/08/17 class:
```{r}
#stop and frisk events by race:
stops_by_race_graph <- ggplot(stops_by_race2, aes(x = race, y = count)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Number of Stop & Frisk Events") +
  xlab("Race") +
  ggtitle("Stop & Frisk Events by Race") +
  geom_text(aes(label = count), vjust = 0, colour = "black")

stops_by_race2

# new race categories in pop file:
# group pop races as such to match the stop and frisk race categories:
## "tp_One race_American Indian and Alaska Native" - row 31
## "tp_One race_Asian" (row 36) + "tp_One race_Native Hawaiian and Other Pacific Islander" (row 44) = one category
## "tp_One race_Black or African American" row 30
## "tp_One race_White" row 29
## "tp_HISPANIC OR LATINO AND RACE_Not Hispanic or Latino_White alone" (row 67) footnote: (b) Hispanics may be of any race, so also are included in applicable race categories


demographic2 <- data.frame(subset(pop[c(31, 36, 44, 30, 29, 67), 1:2]))

typeof(demographic2$Estimate)
str(demographic2)
demographic2$Estimate <- as.numeric(gsub(",", "", as.character(demographic2$Estimate)))
demographic2

demographic2[ , 1]
demographic2$Subject <- c("American Indian and Alaska Native",
              "Asian",
              "Native Hawaiian and Other Pacific Islander",
              "Black or African American",
              "White",
              "White - Hispanic")

#combine "Asian" and "Native Hawaiian and Other Pacific Islander" to match races in sfd file:
52457 + 666
demographic3 <- data.frame(Subject = c("Asian/Native Hawaiian and Other Pacific Islander"),
                           Estimate = 53123)
demographic4 <- demographic2[-c(2:3) , 1:2]
demographic4
demographic4[5, 1:2] <- demographic3
demographic4
# demographic4[5, 2] <- demographic3[c("Asian/Native Hawaiian and Other Pacific Islander"), c("53123")]
demographic4[5, 1] <- "Asian/Native Hawaiian and Other Pacific Islander"
demographic4

population_race_graph <- ggplot(demographic4, aes(x = Subject, y = Estimate)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Race") +
  ylab("Population Estimate") +
  ggtitle("2015 Bronx Population by Race") +
  geom_text(aes(label = Estimate), vjust = 0.05, colour = "black")

#percentage comparison stop and frisk events / pop race cat:
stops_by_race2
demographic4

AIAN <- round(((76 / 7980) * 100), digits = 2) # "American Indian / Alaskan Native"
API <- round(((1043 /	53123) * 100), digits = 2) # "Asian / Pacific Islander" 
BLK <- round(((12919	/ 475378) * 100), digits = 2) # "Black" 
WHT <- round(((2389 / 299869) * 100), digits = 2) # "White" 
WH <- round(((4907 / 146928) * 100), digits = 2) # "White - Hispanic"

#MAKE THIS INTO A MATRIX AND WRITE UP A QUICK SUMMARY FOR WEDNESDAY 
comparison <- matrix(c("Race", "% of each race who were stopped and frisked",
         "American Indian / Alaskan Native", AIAN,
         "Asian / Pacific Islander", API,
         "Black", BLK,
         "White", WHT,
         "White - Hispanic", WH), 
       nrow = , 
       ncol = 2, 
       byrow = TRUE)

comparison2 <- as.data.frame(comparison)

comparison2

comparison2 <- comparison2[-1, 1:2]

comparison2

comparison_graph <- ggplot(comparison2, aes(x = V1, y = V2)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Race") +
  ylab("Percentage") +
  ggtitle("Percentage of Each Race of the Bronx Population Being Stopped & Frisked") +
  geom_text(aes(label = V2), vjust = 0, colour = "black")
```

__03/08/17 Class Update for Team 3:__

  1) Gathered demographic data for the Bronx and compared it to stop and frisk
  events by race.  NYPD crime data were also gathered and will be cleaned within the coming
  weeks to understand the relationship between the reasons police officers  
  stop and frisk versus population crimes in the Bronx by precinct (to pinpoint location).
  Next, we will start to build our forecasting models for probability of being stopped
  and frisked in the Bronx by race, use of force, gender, and other potential variables. 
  
  2) The output (graphs) for the week show the breakdown of 5 races by population
  in the Bronx, stop and frisk events by race, and stop and frisk events by race as a percentage
  of each race within the Bronx population, respectively.  All data are for the Bronx in 2015.

```{r}
population_race_graph
stops_by_race_graph
comparison_graph
```

# Crime Stats for the Bronx: general:

# Clean the data:

https://stackoverflow.com/questions/37509886/how-to-read-merged-excel-cells-with-r 

```{r}
rm(list = ls(all = TRUE))   # cleans everything in the workspace

path <- "/Users/StephanieLangeland/Desktop/Columbia/Applied Data Science/Git/QMSS_G5069_Applied_D_S/Data:Code Book/Raw:Outdated/crime data_combined_raw2.csv"

raw_crime <- read.csv(path)
raw_crime1 <- raw_crime ## create raw_crime1 to alter the data so that raw_crime stays in its original form

raw_crime1 <- raw_crime1[!grepl("TOTAL", raw_crime1$CRIME),] ## remove the rows with totals

raw_crime_2015 <- raw_crime1 ## create a dataset to store only 2015 data

raw_crime_2015 <- raw_crime_2015[ , -c(4:18)] ## remove all years that are not 2015

colnames(raw_crime_2015)[4] <- "crimes_count" ## since all data are form 2015, rename raw_crime_2015$X2015

colnames(raw_crime_2015)[3] <- "crime" ## rename raw_crime_2015$CRIME with lowercase column heading for consistency

colnames(raw_crime_2015)[2] <- "pct" ## rename raw_crime_2015$PCT with lowercase column heading for consistency

cleaned_2015_crime_data <- write.csv(raw_crime_2015, file = "cleaned_2015_crime_data.csv")
```


