---
Title: 'Data Challenge #1'
Authors: "Brandon Wolff, Zachary Heinemann, and Stephanie Langeland	"
Due Date: "02/22/2017"
output: html_document
---

With this new dataset, please address the following:

#1 Can you replicate the 86.1% number? the overall lethality ratio? The ratios for the 
Federal Police, Navy and Army?

```{r}

```

#2 Provide a visualization that presents this information neatly.

```{r}

```

#3 Please show the exact computations you used to calculate them (most likely than not, 
you'll need to do some additional munging in the data to get there).

```{r}

```

#4 If you could not replicate them, please show why and the difference relative to 
your own computations (also, include a neat graph that summarizes this).

```{r}

```

#5 Be very explicit: What are you assuming to generate these computations?
 
  *

Now you know the data more intimately. Think a little bit more about it, and 
answer the following questions:

#6 Is this the right metric to look at? Why or why not?
  
  *

#7 What is the "lethality index" showing explicitly? What is it not showing? What 
is the definition assuming?

  *

#8 With the same available data, can you think of an alternative way to capture the 
same construct? Is it "better"?
 
  *

#9 What additional information would you need to better understand the data?

  *

#10 What additional information could help you better capture the construct behind 
the "lethality index"?

  *
