---
title: 'Data Challenge #1'
author: "Team 3"
date: "2/15/2017"
output: html_document
---

Data Challenge 1 (due on 2/22 at 6PM)

With this new dataset, please address the following:

Can you replicate the 86.1% number? the overall lethality ratio? the ratios for the 
Federal Police, Navy and Army?

```{r}

```

Provide a visualization that presents this information neatly.

```{r}

```

Please show the exact computations you used to calculate them (most likely than not, 
you'll need to do some additional munging in the data to get there).

```{r}

```

If you could not replicate them, please show why and the difference relative to 
your own computations (also, include a neat graph that summarizes this).

```{r}

```

Be very explicit: What are you assuming to generate these computations?
  * 

Now you know the data more intimately. Think a little bit more about it, and 
answer the following questions:

Is this the right metric to look at? Why or why not?
  *

What is the "lethality index" showing explicitly? What is it not showing? What 
is the definition assuming?
  *

With the same available data, can you think of an alternative way to capture the 
same construct? Is it "better"?
  *

What additional information would you need to better understand the data?
  *

What additional information could help you better capture the construct behind 
the "lethality index"?
  *

This is a team assignment. Please create a file on your team GitHub repo where you answer the challenge, including links to your code, graphs.