---
title: "Logistic Regression"
author: "Brandon Wolff"
date: "March 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bronx Logit
```{r}
stop_frisk <- read.csv("2015_stopandfrisk_CLEAN_w_counties.CSV")

sf_bronx <- stop_frisk[stop_frisk$city == 'Bronx',]
```





```{r}
sf_bronx$race <- as.factor(sf_bronx$race)
sf_bronx <- within(sf_bronx, race <- relevel(race, ref = 4))

summary(fit <- glm(pforce ~ race + sex, data = sf_bronx, family = "binomial"))
```

When running a logistic regression for the Bronx only on "pforce" meaning the police used force or did not with race 
(white as the reference group) and sex alone we see that this model does not have significant coefficients for any of
the minority groups except American Indian / Alaskan Native people. According to the results being American Indian / 
Alaskan Native increases the log odds relative to being white and net of all other factors by 1.72 and this is significant
at the .05 level.

```{r}
summary(fit <- glm(pforce ~ race + sex + build+ ac_incid + ac_time + armed + offunif + inout, data = sf_bronx, family = binomial(link = "logit")))
```


```{r}
summary(fit <- glm(pforce ~ race + sex + build+ ac_incid + ac_time + armed + offunif + inout + arstmade + explnstp + frisked, data = sf_bronx, family = binomial(link = "logit")))
```

When running a logistic regression for the Bronx on "pforce" meaning the police used force or did not with race 
(white as the reference group), Sex, body build, "ac_incid" which is whther the event took place in a high crime
or not, "ac_time" (if event took place at abnormal time), "armed" (person was armed or not), "offunif" (if office
was in uniform or not), "inout" (whether event took place in or outside), "arstmade" (if arrest was made or not),
"explnstp" (if the person was given reason for the stop or not), and finally "frisked" (if the person was frisked
or not) we see that this model does not have significant coefficients for any of the minority groups except American
Indian / Alaskan Native people. According to the results being American Indian / Alaskan Native increases the log odds
relative to being white and net of all other factors by 1.90 and this is significant at the .05 level. It is also good
to note that according to the results having a muscular build increases the log odds relative to being heavy and net of
all other factors by 0.85 and this is significant at the .01 level.