# #######################################################################
# File-Name:  demographics and crime stats_v1
# Version:  1
# Date: 03/04/17 
# Author: Stephanie Langeland 
# Purpose: Find demographic and crime stats data
# Input Files: 2015_stopandfrisk_CLEAN_w_counties.csv 
# Output Files: None yet
# Data Output: None
# Dependencies: ?
# Required by: ?
# Status: In progress 
# Machine: Stephanie's 2011 MacBook Pro
# R version: 3.3.1
# #######################################################################

Stop and Frisk Data: General:
```{r}
rm(list = ls(all = TRUE))   # cleans everything in the workspace

sfd <- read.csv("/Users/StephanieLangeland/Desktop/Columbia/Applied Data Science/Git/QMSS_G5069_Applied_D_S/Data:Code Book/Cleaned/2015_stopandfrisk_CLEAN_w_counties.csv") # cleaned stop and frisk data 

# SPACE FOR SOMEONE ELSE TO COMMENT OUT MY PATH ABOVE AND PUT THEIR PATH IN HERE
```

Stop and Frisk Data: Time period of the data:
```{r}
# convert to date format:
library(lubridate)

summary(sfd$datestop)
summary(is.na(sfd$datestop))
typeof(sfd$datestop)
sfd$datestop <- mdy(sfd[ , 5]) # column 5 is the DATE OF STOP (M-D-YYYY) <- the code book is wrong, it says that the format is (MM-DD-YYY)
summary(is.na(sfd$datestop))

min(sfd$datestop) #confirmed that the data are complete for 2015
max(sfd$datestop)

summary((sfd$dob)) # SUSPECT'S DATE OF BIRTH (CCYY-MM-DD) is all missing -> use `age` variable instead
```

Demographic data for the Bronx: General 
data source: https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF 
```{r}
#the raw dataset was such a mess that I had to clean it in excel - the follwing is the cleaned version:

pop <- read.csv("/Users/StephanieLangeland/Desktop/Columbia/Applied Data Science/Git/QMSS_G5069_Applied_D_S/Data:Code Book/Cleaned/CLEAN_ACS_15_5YR_DP05.csv") 
```
#stop and frisk by race per capita sort of thing
#figure out the proportion of each race being stopped and frisk for this week's presentation

Crime Stats for the Bronx:
#does probable cause match the popular crimes in the area?
#stop and frisk by crime per capita sort of thing
```{r}

```




